---
title: "ID in basketball"
author: 
- "Edgar Santos Fernandez"
- "Francesco Denti"
- "Kerrie Mengersen"
- "Antonietta Mira"
date: "2/6/2020"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
This file contains intructions to reproduce some of the results in the article
"The role of intrinsic dimension in high-resolution player tracking data - Insights in basketball".



```{r, eval = T, echo =F}
.libPaths( c( "~/R_LIBS" , .libPaths() ) )
setwd("~/Desktop/Research/VRES")
```

```{r, eval = T, message=F, warning=F}
library(data.table)
library(Rcpp)
library(RcppArmadillo)
library(farver)
library(ggplot2)
library(mvtnorm)
library(dplyr)
library(reshape2)
library(plot3D)
library(Matrix)
library(superheat)
library(mcclust)
library(RCurl)
library(jsonlite)
library(sp)
library(zoo)
library(lubridate)
library(stringr)
library(gridExtra)
library(grid)
library(RColorBrewer)
library(ggrepel)
library(viridis)
library(tidyr)

```

```{r, eval = T, message=F, warning=F}
# TODO:
# - Finish visualisation of player movements over time
# - Characterisation of players position in space
# - Adapt script to automatically take NBA dataset, create gifs of each play, store hidalgo stuff, 
#   and other properties of the players in the play
# - Create aggregate statistics about the game - points scored, where they were scored, by whom, path to generate a shot
```
Hidalgo is implemented in C++ and the sampler and other functions are located in the file ```Hidalgo_Code.cpp```. 
Other useful functions from files ```Helpers_Code.R```  and ```plot functions.R``` need also to be sourced. 


```{r, eval = T, message=F, warning=F}
Rcpp::sourceCpp("./Code/Hidalgo_Code.cpp")
source("./Code/Helpers_Code.R")
source('./Code/plot functions.R')
```

The most of the movement data from the game is on the file ```CLEatGSW.RDS```.

```{r, eval = T}
allm <- readRDS("CLEatGSW.RDS")
allm$descrip <- ifelse(!is.na(allm$HOMEDESCRIPTION), as.character(allm$HOMEDESCRIPTION), as.character(allm$VISITORDESCRIPTION))

# selecting shot made and missed
allm <- dplyr::filter(allm, EVENTMSGTYPE %in% c(1,2) )


select_id <- function(all.movements = all.movements, id = id){
  e1 <- dplyr::filter(all.movements, event.id == id) 
  e1$x_loc <- as.numeric(as.character(e1$x_loc))
  e1$y_loc <- as.numeric(as.character(e1$y_loc))
  e1$shot_clock <- as.numeric(as.character(e1$shot_clock))
  e1$shot_clock2 <- round(e1$shot_clock,1)
  e1 <- dplyr::arrange(e1, datetime2)
  
  e1$ii <- 1
  e1 <- e1 %>% dplyr::mutate(act = ifelse(shot_clock2 > lag(shot_clock2), ii + 1,ii)) # need to split actions because multiple actions are mingled
  
  breaks <- c(ymd_hms('2010-01-01 00:00:00'), e1[which(e1$act == 2),]$datetime2, Sys.time())
  e1$action <- cut.POSIXt(e1$datetime2, breaks = breaks, labels = 1:(length(breaks)-1))
  
  e1 <- dplyr::filter(e1, shot_clock2 < 24)
  e1 <- e1 %>% dplyr::group_by(event.id) %>%  dplyr::mutate(ids = dplyr::row_number())
  
}


# table with the action to be used.
# this was done manually because many events have issues with the clock
# 
table_fixing_clock <- read.csv('across plays fixing clock issues.csv')
table_fixing_clock$quarter <- factor(table_fixing_clock$quarter)
table_fixing_clock$EVENTMSGTYPE <- factor(table_fixing_clock$EVENTMSGTYPE)

allm$EVENTMSGTYPE <- droplevels(allm$EVENTMSGTYPE)
allm$ball_possession <- factor(allm$ball_possession)

allm <- allm %>% left_join(table_fixing_clock, by = c('event.id', 'quarter', 'EVENTMSGTYPE', 'ball_possession'))

# will exclude the plays with issues according to the visualization done with assess

allm <- dplyr::filter(allm, comment != 'e')

```

```{r, eval = T}
unique_events <- allm %>% 
                 dplyr::select(event.id, EVENTMSGTYPE, ball_possession, quarter, descrip, SCORE) %>% 
                 distinct() %>%
                 mutate(scored_on_play = case_when(is.na(SCORE) ~ 0,
                                                   TRUE ~ 1))

ev <- unique(allm$event.id) # events


# computing ID for event 6 
#i = 7
i = 1

s15 <- select_id(all.movements = allm, id = ev[i]) 
s15 <- filter(s15, action == table_fixing_clock[table_fixing_clock$event.id == ev[i],]$action)

if(unique(s15$quarter) %in% c(1,2) ){
  if(unique(s15$ball_possession) == 'home') { 
         s15 <- filter(s15, datetime2 > min(filter(s15, lastname == 'ball', x_loc < 47)$datetime2))
         }
if(unique(s15$ball_possession) == 'away') {
         s15 <- filter(s15, datetime2 > min(filter(s15, lastname == 'ball', x_loc > 47)$datetime2))
       }
    }


#filter offensive court only
if(unique(s15$quarter) %in% c(3,4) ){
      if(unique(s15$ball_possession) == 'home') { 
        s15 <- filter(s15, datetime2 > min(filter(s15, lastname == 'ball', x_loc > 47)$datetime2))
      }
if(unique(s15$ball_possession) == 'away') {
        s15 <- filter(s15, datetime2 > min(filter(s15, lastname == 'ball', x_loc < 47)$datetime2))
      }
}


```


```{r, eval = T}

stamps <- unique(s15$datetime2)[seq(1, length(unique(s15$datetime2)), by = 10)] 
unique(s15$lastname)

# Save stamps for later
stampsSave <- data.frame(stamps)

s15 <- filter(s15, datetime2 %in% stamps) %>% dplyr::arrange(event.id, datetime2, team, lastname)
  
s15$team <- as.character(s15$team)
  #levels(s15$team) <- c('away', 'home', 'ball')
s15$team <- ifelse(is.na(s15$team) , 'ball', s15$team)
####################

  s15$x <- s15$x_loc
  s15$y <- s15$y_loc
  s15$player <-  rep(c(paste0('a', 1:5), paste0('h', 1:5), 'ball') , times = round(nrow(s15)/11))[1:nrow(s15)]

  s15 <- s15 %>% arrange(player, datetime2)
  s15$stamp <- rep(1:length(stamps),times = 11)
  
  s15 <- s15 %>% dplyr::group_by(player, lastname) %>% 
    dplyr::mutate(dist = sqrt( (x-lag(x)) ^ 2 + (y - lag(y)) ^ 2) ) #%>% dplyr::group_by(player, shot_clock) 

  if(unique(s15$ball_possession) == 'home') { 
     players <-  c("ball", unique(s15[s15$team == 'home',]$lastname))}
  
  if(unique(s15$ball_possession) == 'away') { 
    players <-  c("ball", unique(s15[s15$team == 'away',]$lastname))}
  
  if(unique(s15$ball_possession) == 'home') { 
    players2 <-  c(unique(s15[s15$team == 'away',]$lastname))}
  
  if(unique(s15$ball_possession) == 'away') { 
    players2 <-  c(unique(s15[s15$team == 'home',]$lastname))}
  
s15$lastname <- factor(s15$lastname)

colp <- brewer.pal(12, "Paired")[c(6, 2,4,8, 10,12)]

# plotting the movement of the players on attack
# https://youtu.be/jb57MFQLoRo?t=17

# Plot for the home team with the ball
p1 <- P_180 + 
      geom_point(data = dplyr::filter(s15,lastname %in% players ), aes(x_loc, -y_loc, group = lastname, col = lastname, shape = team), size =1 ) + 
      geom_path(data = dplyr::filter(s15,lastname %in% players), aes(x_loc, -y_loc, group = lastname, col = lastname), size = 0.3) +
      geom_text(data = dplyr::filter(s15,lastname %in% players), aes(x_loc, -y_loc+1, label = stamp, col = lastname), size = 1.5)+
      scale_color_manual(values = colp) +
      theme_bw() + ggtitle(paste0('event_id = ', unique(s15$event.id))) 

# Plot for the other team without ball
p1_other <- P_180 + 
            geom_point(data = dplyr::filter(s15,lastname %in% players2), aes(x_loc, -y_loc, group = lastname, col = lastname, shape = team), size =1 ) + 
            geom_path(data = dplyr::filter(s15,lastname %in% players2), aes(x_loc, -y_loc, group = lastname, col = lastname), size = 0.3) +
            geom_text(data = dplyr::filter(s15,lastname %in% players2), aes(x_loc, -y_loc+1, label = stamp, col = lastname), size = 1.5)+
            scale_color_manual(values = colp) +
            theme_bw() + ggtitle(paste0('event_id = ', unique(s15$event.id))) 

# Visualise the home and away teams
grid.arrange(p1, p2, nrow = 2, ncol = 1, heights=c(10,5), widths=c(11))

dev.copy(jpeg,filename="Temp - Offense.jpg");
dev.off ();

grid.arrange(p1_other, p2, nrow = 2, ncol = 1, heights=c(10,5), widths=c(11))

dev.copy(jpeg,filename="Temp - Defense.jpg");
dev.off ();


```
  
```{r, eval = T, cache = T}
dist_stamp <- s15 %>% dplyr::group_by(datetime2) %>%
    dplyr::summarize(dist = sum(dist, na.rm = T))

curryPosition <- s15 %>%
                 subset(firstname == "Stephen")

kyriePosition <- s15 %>%
                 subset(firstname == "Kyrie")

klovePosition <- s15 %>%
                 subset(firstname == "Kevin")
  
s15_ <- data.table::dcast(setDT(s15), event.id + quarter + EVENTMSGTYPE+
                EVENTMSGACTIONTYPE + SCORE + SCOREMARGIN + HOMEDESCRIPTION +
                VISITORDESCRIPTION + datetime2 + shot_clock + game_clock ~ player, value.var=c("x", "y", "radius"))
  
s15_ <- s15_ %>% dplyr::select(-c("radius_a1","radius_a2","radius_a3", "radius_a4","radius_a5","radius_h1",        
                                "radius_h2","radius_h3","radius_h4","radius_h5") )

s15_ <- as.data.frame(s15_)
  
#vars <- c(names(s15_)[grepl('_h', names(s15_))] , names(s15_)[grepl('_a', names(s15_))], "x_ball", "y_ball")  
vars <- c(names(s15_)[grepl('_h', names(s15_))] , names(s15_)[grepl('_a', names(s15_))])  

DATA <- s15_[, vars] 
```

Applying HIDALGO to the movement data with dimension D = 20. 
```{r, eval = T, cache = T}
## Declare Hidalgo Parameters
nsim = 20000 # simulations
burn_in = 10000 # burn in period
L      <- 3 
Lista <- data.preprocessing(DATA, q = 3) # computes the mus

# Apply Hidalgo
out3R   <- HIDALGO(mu_obser = Lista[[1]],
                       L = L,
                       NSIM = nsim, 
                       burn_in = burn_in,
                       thinning = 1, 
                       verbose = F, verbose_step = 1000,
                       a0_d = 2, b0_d = 1, Nq = Lista[[2]],
                       q = Lista[[3]],CSI = .75, 
                       t1 = .5, t2 = .001,REPULSIVE = T,
                       alpha = rep(1,L),D = 20,TRUNC = F)
  
tracked_d3R=Tracking(out3R$AC,out3R$AD, N = nrow(DATA), NSIM = nsim)

ID <- apply(tracked_d3R,1,median)
```

Plotting the ID accross time stamps:

```{r, eval = T, cache = T}    
df15 <- data.frame(id = ID, i = 1:length(ID))
 
p2 <- ggplot(df15) + 
  geom_line(aes(y = id, x = i, col = id), size = 1.5)+
  scale_color_viridis()+
  xlab('frame')+ ylab('ID')+
  theme_bw() 

grid.arrange(p1, p2, nrow = 2, ncol = 1, heights=c(10,5), widths=c(11))

dev.copy(jpeg,filename="Temp - Offense.jpg");
dev.off ();

grid.arrange(p1_other, p2, nrow = 2, ncol = 1, heights=c(10,5), widths=c(11))

dev.copy(jpeg,filename="Temp - Defense.jpg");
dev.off ();

```


```{r, eval = T}
grid.arrange(p1, p2, nrow = 2, ncol = 1, heights=c(10,5), widths=c(11))
```

Heatmap of the posterior similarity matrix of the play:

```{r, eval = T}
M_away <- mcclust::comp.psm(t(out3R$AC))
set.seed(2020)
superheat::superheat(M_away, 
                            smooth.heat = T,
                            legend = T,
                            bottom.label.size = .15,
                            left.label.size = .15,
                            bottom.label.text.size = 3,
                            left.label.text.size = 3,
                            left.label.text.angle = 0,
                            bottom.label.text.angle = 90 )

```
Try and say something spatially about where each player is relative to the other players:

```{r, eval = T}

# Attach back the intrinsic dimension to the main data
# Get the rate of change of distance - help visualise when a pass has occurred 


newData <- s15_ %>% 
           cbind(ID) %>%
           mutate(seconds_elapsed = as.numeric(abs(shot_clock - lag(shot_clock))),
                  # Player velocities
                  ball_velocity = sqrt((x_ball-lag(x_ball))^2 + (y_ball-lag(y_ball))^2)/seconds_elapsed,
                  h1_velocity = sqrt((x_h1-lag(x_h1))^2 + (y_h1-lag(y_h1))^2)/seconds_elapsed,
                  h2_velocity = sqrt((x_h2-lag(x_h2))^2 + (y_h2-lag(y_h2))^2)/seconds_elapsed,
                  h3_velocity = sqrt((x_h3-lag(x_h3))^2 + (y_h3-lag(y_h3))^2)/seconds_elapsed,
                  h4_velocity = sqrt((x_h4-lag(x_h4))^2 + (y_h4-lag(y_h4))^2)/seconds_elapsed,
                  h5_velocity = sqrt((x_h5-lag(x_h5))^2 + (y_h5-lag(y_h5))^2)/seconds_elapsed,
                  a1_velocity = sqrt((x_a1-lag(x_a1))^2 + (y_a1-lag(y_a1))^2)/seconds_elapsed,
                  a2_velocity = sqrt((x_a2-lag(x_a2))^2 + (y_a2-lag(y_a2))^2)/seconds_elapsed,
                  a3_velocity = sqrt((x_a3-lag(x_a3))^2 + (y_a3-lag(y_a3))^2)/seconds_elapsed,
                  a4_velocity = sqrt((x_a4-lag(x_a4))^2 + (y_a4-lag(y_a4))^2)/seconds_elapsed,
                  a5_velocity = sqrt((x_a5-lag(x_a5))^2 + (y_a5-lag(y_a5))^2)/seconds_elapsed)

player_distance_matrix <- newData %>%
                          select(x_a1:y_h5) 

ball_matrix <- player_distance_matrix %>%
               select(subset = c(x_ball,
                                 y_ball))

player_distance_matrix <- player_distance_matrix %>%
                          select(subset = -c(x_ball,
                                             y_ball))

# Nuke data made in the next step
newData <- newData[, 1:47]

# Hard code net coordinates for now. Need these to characterize how a player moves towards the basket
# e.g. have they broken away from their defender. This is only specific to unique_events row 7. No 
# idea how it stacks up with other plays

basket_x <- 5.2
basket_y <- 25


# Do distance of players from each other
for (i in 1:10) {

    for (j in 1:10) {

        temp_x <- as.data.frame(player_distance_matrix[,i] - player_distance_matrix[,j])
        temp_y <- as.data.frame(player_distance_matrix[,i+10] - player_distance_matrix[,j+10])
        temp_abs <- sqrt(temp_x^2 + temp_y^2)
        
        if (i <= 5) {
            if (j <= 5) {
                colnames(temp_x) <- paste0("x_a", i, "_rel_a",j)
                colnames(temp_y) <- paste0("y_a", i, "_rel_a",j)
                colnames(temp_abs) <- paste0("abs_a", i, "_rel_a",j)
            } else {
                colnames(temp_x) <- paste0("x_a", i, "_rel_h",j-5)
                colnames(temp_y) <- paste0("y_a", i, "_rel_h",j-5)
                colnames(temp_abs) <- paste0("abs_a", i, "_rel_h",j-5)
            }
        } else {
            if (j <= 5) {
                colnames(temp_x) <- paste0("x_h", i-5, "_rel_a",j)
                colnames(temp_y) <- paste0("y_h", i-5, "_rel_a",j)
                colnames(temp_abs) <- paste0("abs_h", i-5, "_rel_a",j)
            } else {
                colnames(temp_x) <- paste0("x_h", i-5, "_rel_h",j-5)
                colnames(temp_y) <- paste0("y_h", i-5, "_rel_h",j-5)
                colnames(temp_abs) <- paste0("abs_h", i-5, "_rel_h",j-5)
            }
        }
        
        newData <- cbind(newData, temp_x)
        newData <- cbind(newData, temp_y)
        newData <- cbind(newData, temp_abs)
    }
  
    # Get Draymond's distance from the basket
    temp_x <- as.data.frame(player_distance_matrix[,i] - ball_matrix[,1])
    temp_y <- as.data.frame(player_distance_matrix[,i+10] - ball_matrix[,2])
    temp_abs <- sqrt(temp_x^2 + temp_y^2)
  
    if (i <= 5) {
        colnames(temp_x) <- paste0("x_h", i-5, "_rel_ball")
        colnames(temp_y) <- paste0("y_h", i-5, "_rel_ball")
        colnames(temp_abs) <- paste0("abs_h", i-5, "_rel_ball")
    } else {
        colnames(temp_x) <- paste0("x_h", i-5, "_rel_ball")
        colnames(temp_y) <- paste0("y_h", i-5, "_rel_ball")
        colnames(temp_abs) <- paste0("abs_h", i-5, "_rel_ball")
    }
    
    newData <- cbind(newData, temp_x)
    newData <- cbind(newData, temp_y)
    newData <- cbind(newData, temp_abs)
    
    # Get Draymond's distance from the ball
    temp_x <- as.data.frame(player_distance_matrix[,i] - basket_x)
    temp_y <- as.data.frame(player_distance_matrix[,i+10] - basket_y)
    temp_abs <- sqrt(temp_x^2 + temp_y^2)
  
    if (i <= 5) {
        colnames(temp_x) <- paste0("x_h", i-5, "_rel_basket")
        colnames(temp_y) <- paste0("y_h", i-5, "_rel_basket")
        colnames(temp_abs) <- paste0("abs_h", i-5, "_rel_basket")
    } else {
        colnames(temp_x) <- paste0("x_h", i-5, "_rel_basket")
        colnames(temp_y) <- paste0("y_h", i-5, "_rel_basket")
        colnames(temp_abs) <- paste0("abs_h", i-5, "_rel_basket")
    }
    
    newData <- cbind(newData, temp_x)
    newData <- cbind(newData, temp_y)
    newData <- cbind(newData, temp_abs)
    
}

# For each player, visualise it's distance from the other players at one time point
# TODO: Make animation, for each player, of their relative distance to the other players.

draymond_x <- newData %>%
              select(contains("x_h3"))

draymond_y <- newData %>%
              select(contains("y_h3"))

bogut_x <- newData %>%
           select(contains("x_h1_"))

bogut_y <- newData %>%
           select(contains("y_h1_"))

draymond_abs <- newData %>%
                select(contains("abs_h3")) 

draymond_abs_2 <- cbind(draymond_abs,
                  as.data.frame(newData$shot_clock)) %>%
                  gather(key="Player", value="Distance", abs_h3_rel_a1:abs_h3_rel_basket) %>%
                  mutate(Player = case_when(Player == "abs_h3_rel_h1" ~ "Green Relative to Bogut",
                                            Player == "abs_h3_rel_h2" ~ "Green Relative to Curry",
                                            Player == "abs_h3_rel_h3" ~ "Green Relative to Green",
                                            Player == "abs_h3_rel_h4" ~ "Green Relative to Rush",
                                            Player == "abs_h3_rel_h5" ~ "Green Relative to Thompson",
                                            Player == "abs_h3_rel_a1" ~ "Green Relative to Irving",
                                            Player == "abs_h3_rel_a2" ~ "Green Relative to James",
                                            Player == "abs_h3_rel_a3" ~ "Green Relative to Love",
                                            Player == "abs_h3_rel_a4" ~ "Green Relative to Mozgov",
                                            Player == "abs_h3_rel_a5" ~ "Green Relative to Smith",
                                            Player == "abs_h3_rel_basket" ~ "Green Relative to Basket",
                                            Player == "abs_h3_rel_ball" ~ "Green Relative to Ball")) 

colnames(draymond_abs_2) <- c("ShotClock", "Player", "Distance")


# Visualise Draymonds distance from other players
draymond_distance <- ggplot(draymond_abs_2) + 
                     geom_line(aes(y = Distance, x = 24 - ShotClock, col=Player, group=Player), size = 1.5) +
                     xlab('Play Time Elapsed (s)') + ylab('Distance from Other Players (m)') +
                     ggtitle("Draymond Distance to Other Players") +
                     theme_bw() 

plot(draymond_distance)

```

Heatmap of the posterior similarity matrix of the play:

```{r, eval = T}
           
newData_T <- newData %>%
             gather(key="Player", value="Velocity", h1_velocity:a5_velocity) %>%
             mutate(Player = case_when(Player == "h1_velocity" ~ "Bogut",
                                       Player == "h2_velocity" ~ "Curry",
                                       Player == "h3_velocity" ~ "Green",
                                       Player == "h4_velocity" ~ "Rush",
                                       Player == "h5_velocity" ~ "Thompson",
                                       Player == "a1_velocity" ~ "Irving",
                                       Player == "a2_velocity" ~ "James",
                                       Player == "a3_velocity" ~ "Love",
                                       Player == "a4_velocity" ~ "Mozgov",
                                       Player == "a5_velocity" ~ "Smith"))



# Plot ID against ball velocity
p3 <- ggplot(newData) + 
      geom_line(aes(y = ball_velocity, x = ID, col = ID), size = 1.5) +
      scale_color_viridis() +
      xlab('ID') + ylab('Ball Velocity')+
      theme_bw() 

p4 <- ggplot(newData_T %>%
      subset(Player == "Bogut")) + 
      geom_line(aes(y = Velocity, x = ID, colour = Player, group=Player), size = 1.5) + 
      xlab('Shot Clock Time') + ylab('Player Velocity')+
      theme_bw()

p5 <- ggplot(newData_T %>%
      subset(Player == "Curry")) + 
      geom_line(aes(y = Velocity, x = ID, colour = Player, group=Player), size = 1.5) + 
      xlab('Shot Clock Time') + ylab('Player Velocity')+
      theme_bw()

p6 <- ggplot(newData_T %>%
      subset(Player == "Green")) + 
      geom_line(aes(y = Velocity, x = ID, colour = Player, group=Player), size = 1.5) + 
      xlab('Shot Clock Time') + ylab('Player Velocity')+
      theme_bw()

p7 <- ggplot(newData_T %>%
      subset(Player == "Rush")) + 
      geom_line(aes(y = Velocity, x = ID, colour = Player, group=Player), size = 1.5) + 
      xlab('Shot Clock Time') + ylab('Player Velocity')+
      theme_bw()

p8 <- ggplot(newData_T %>%
      subset(Player == "Thompson")) + 
      geom_line(aes(y = Velocity, x = ID, colour = Player, group=Player), size = 1.5) + 
      xlab('Shot Clock Time') + ylab('Player Velocity')+
      theme_bw()

plot(p3)
plot(p4)
plot(p5)
plot(p6)
plot(p7)
plot(p8)


```

Spread data back out and see if the velocities at certain time points are correlated:

```{r, eval = T}



```